---
title: "Geocoded Unified BMF Quality Report"
format:
  html:
    theme: cosmo
    embed-resources: true
params:
  report_data_path: ""
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressPackageStartupMessages(library(knitr))
if (!file.exists(params$report_data_path)) stop("Report data file not found: ", params$report_data_path)
r <- readRDS(params$report_data_path)
fmt_int <- function(x) format(as.integer(x), big.mark = ",")
pct <- function(a, b) sprintf("%.2f %%", 100 * as.numeric(a) / as.numeric(b))
```

::: {.callout-note appearance="simple"}
**Generated:** `r r$timestamp`
**Data set:** `geocoding/unified-bmf/latest/bmf_unified_geocoded.parquet` (the Unified BMF with coordinates attached)
:::

# What this report covers

The geocoded Unified BMF has exactly the rows of the Unified BMF plus
location columns. Coordinates are attached by matching each organization's
mailing address against the Urban Institute geocoder. Rows without a usable
mailing address (for example, no street and no city) cannot be sent to the
geocoder and keep empty location columns.

# Coverage

```{r}
tbl <- data.frame(
  Measure = c("Unified BMF rows (one per EIN)",
              "Rows with a usable mailing address (sent to the geocoder)",
              "Distinct addresses geocoded",
              "Rows with latitude and longitude",
              "Share of all rows with coordinates",
              "Share of geocodable rows with coordinates",
              "Geocoder match rate on distinct addresses"),
  Value = c(fmt_int(r$total_master_rows),
            fmt_int(r$geocodable_rows),
            fmt_int(r$unique_addresses_geocoded),
            fmt_int(r$rows_with_lat_lon),
            pct(r$rows_with_lat_lon, r$total_master_rows),
            pct(r$rows_with_lat_lon, r$geocodable_rows),
            sprintf("%.2f %%", as.numeric(r$geocoder_match_rate_pct))),
  stringsAsFactors = FALSE)
kable(tbl, format = "html")
```

# How to read it

- Rows without coordinates are almost all rows without a usable mailing
  address, not geocoder failures. The gap between "rows" and "rows with a
  usable mailing address" is that group.
- The geocoder returns a match score per address; the published columns
  carry the score, match type and address type so users can apply their own
  threshold. This report counts any returned coordinate as geocoded.
- For column-level completeness of the non-location columns, see the
  Unified BMF quality report; the geocoded file inherits them unchanged.
