---
title: "Unified BMF Quality Report"
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    embed-resources: true
    page-layout: full
params:
  report_data_path: ""
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
suppressPackageStartupMessages({
  library(data.table)
  library(jsonlite)
  library(knitr)
})

if (!file.exists(params$report_data_path)) {
  stop("Report data file not found: ", params$report_data_path)
}
report <- readRDS(params$report_data_path)
fmt_int <- function(x) format(as.integer(x), big.mark = ",")
fmt_pct <- function(x) sprintf("%.2f %%", x)
```

::: {.callout-note appearance="simple"}
**Generated:** `r report$timestamp`
**Quality gate:** `r if (isTRUE(report$passed)) "**PASSED**" else "**FAILED**"`
**Total rows (unique EINs):** `r fmt_int(report$total_rows)`
**Source files stacked:** `r report$n_input_files`
**Vintages contributing rows:** `r report$n_vintages`
:::

# Summary

The Unified BMF is a single-row-per-EIN consolidation drawn from the
most-recent vintage in which each EIN appears across both the
current monthly BMF pipeline and the legacy 501CX-NONPROFIT-PX
pipeline. See [chapter 11](https://urbaninstitute.github.io/nccs-data-bmf/11-master-bmf.html)
for the full design.

| Metric | Value |
|--------|-------|
| Total rows | `r fmt_int(report$total_rows)` |
| Distinct EINs | `r fmt_int(report$distinct_eins)` |
| Rows == distinct EINs (uniqueness gate) | `r if (isTRUE(report$ein_unique)) "✅ yes" else "❌ NO"` |
| Input CSV files | `r report$n_input_files` |
| Vintages with at least one surviving row | `r report$n_vintages` |

# Rows by source pipeline

```{r}
if (!is.null(report$rows_by_source) && length(report$rows_by_source$bmf_source) > 0) {
  src_dt <- data.table(
    bmf_source = report$rows_by_source$bmf_source,
    n_rows     = report$rows_by_source$n_rows
  )
  src_dt[, pct := round(100 * n_rows / sum(n_rows), 2)]
  kable(src_dt, format = "html",
        col.names = c("Source", "Rows", "Percent"),
        format.args = list(big.mark = ","))
}
```

# First / last year coverage

```{r}
yr <- report$year_range
yr_dt <- data.table(
  metric = c("Earliest first_year_in_bmf",
             "Latest first_year_in_bmf",
             "Earliest last_year_in_bmf",
             "Latest last_year_in_bmf"),
  value  = c(yr$min_first_year, yr$max_first_year,
             yr$min_last_year,  yr$max_last_year)
)
kable(yr_dt, format = "html",
      col.names = c("Metric", "Year"))
```

# Vintages observed per EIN

How many distinct BMF vintages did each EIN appear in?

```{r}
vo <- report$vintages_observed
vo_dt <- data.table(
  metric = c("Minimum",
             "Average",
             "Maximum",
             "EINs that appear in only one vintage"),
  value  = c(fmt_int(vo$min_vintages),
             sprintf("%.1f", as.numeric(vo$avg_vintages)),
             fmt_int(vo$max_vintages),
             fmt_int(vo$n_only_once))
)
kable(vo_dt, format = "html",
      col.names = c("Metric", "Value"))
```

# Vintage histogram

Number of surviving Unified BMF rows contributed by each vintage. A
vintage with many rows contributed many EINs whose latest sighting
was in that month.

```{r}
hist_df <- report$vintage_histogram
if (!is.null(hist_df) && nrow(hist_df) > 0) {
  hist_dt <- as.data.table(hist_df)
  setorder(hist_dt, -n_rows)
  kable(hist_dt, format = "html",
        col.names = c("Vintage (YYYY-MM)", "Source", "Rows contributed"),
        format.args = list(big.mark = ","))
}
```

# Column completeness

Top 30 most-complete columns (highest non-null fraction):

```{r}
# The completeness table carries an overall non-null share plus one share per
# source (current-monthly vs legacy rows). Label every column it has; a
# mismatch between the table and col.names aborts kable() and with it the
# whole render (that is what silently stopped the HTML on 2026-09-15).
comp <- as.data.table(report$completeness)
comp_cols <- intersect(c("column", "pct_non_null", "pct_current", "pct_legacy"),
                       names(comp))
comp <- comp[, ..comp_cols]
comp_names <- c(column = "Column", pct_non_null = "Non-null %",
                pct_current = "Current rows %", pct_legacy = "Legacy rows %")[comp_cols]
setorder(comp, -pct_non_null)
kable(head(comp, 30), format = "html", col.names = unname(comp_names))
```

Bottom 30 least-complete columns:

```{r}
setorder(comp, pct_non_null)
kable(head(comp, 30), format = "html", col.names = unname(comp_names))
```

# Columns populated by only one source

Columns that are well populated in one pipeline's rows and empty in the
other's. These are expected where a field exists in only one source (for
example fields the legacy extracts never carried), and worth a look where
it does not.

```{r}
flags <- report$source_completeness_flags
if (!is.null(flags) && nrow(flags) > 0) {
  flags <- as.data.table(flags)[, ..comp_cols]
  kable(flags, format = "html", col.names = unname(comp_names))
} else {
  cat("None.")
}
```

# Notes

- **All output columns are VARCHAR.** The master build forces every
  column to string type so DuckDB's `UNION ALL BY NAME` can combine
  the legacy slim per-vintage schema with the current full schema
  without per-column type-mismatch errors. Cast in your downstream
  tool: `as.numeric(asset_amount)` in R,
  `pd.to_numeric(df['asset_amount'], errors='coerce')` in pandas, etc.
  Lat/lon in the geocoded master output are explicitly cast to
  DOUBLE.
- **Legacy-only EINs have sparse columns.** EINs whose latest
  appearance is in a pre-2014 legacy vintage carry only the legacy
  slim schema; current-pipeline-only columns (e.g. NTEEv2 fields,
  modern address quality flags) are NULL. The `bmf_source` column
  flags the source of each surviving row.
