---
title: "BMF Pipeline Quality Report"
subtitle: "Post-Transformation Validation Results"
date: "`r Sys.time()`"
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    code-fold: true
    self-contained: true
  pdf:
    toc: true
    toc-depth: 2
params:
  report_data_path: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)

library(ggplot2)
library(data.table)
library(scales)

# Load report data from RDS file
if (!is.null(params$report_data_path) && file.exists(params$report_data_path)) {
  report <- readRDS(params$report_data_path)
} else {
  stop("Report data not found. Please provide a valid report_data_path parameter.")
}
```

# Executive Summary

```{r summary-badge}
status_color <- if (report$passed) "#28a745" else "#dc3545"
status_text <- if (report$passed) "PASSED" else "FAILED"
```

::: {.callout-note appearance="simple" style="background-color: `r status_color`20; border-left-color: `r status_color`;"}
## Overall Status: `r status_text`

**Timestamp:** `r format(report$timestamp, "%Y-%m-%d %H:%M:%S")`
:::

```{r summary-table}
summary_df <- data.frame(
  Metric = c(
    "Final Row Count",
    "Final Column Count",
    "Row Preservation",
    "Missing Expected Columns",
    "Unique EINs"
  ),
  Value = c(
    format(report$row_count, big.mark = ","),
    as.character(report$column_count),
    if (report$row_preservation) "Yes" else "No",
    as.character(length(report$missing_columns)),
    if (!is.null(report$summary_stats$unique_eins))
      format(report$summary_stats$unique_eins, big.mark = ",") else "N/A"
  )
)

knitr::kable(summary_df, col.names = c("Metric", "Value"), align = c("l", "r"))
```

```{r missing-cols, results='asis'}
if (length(report$missing_columns) > 0) {
  cat("\n### Missing Expected Columns\n\n")
  cat("The following columns were expected but not found:\n\n")
  for (col in report$missing_columns) {
    cat(sprintf("- `%s`\n", col))
  }
}
```

```{r critical-issues, results='asis'}
if (length(report$critical_field_issues) > 0) {
  cat("\n### Critical Field Issues\n\n")
  cat("::: {.callout-warning}\n")
  cat("The following critical fields have NULL values:\n\n")
  for (field in names(report$critical_field_issues)) {
    cat(sprintf("- **%s**: %s NULL values\n",
                field,
                format(report$critical_field_issues[[field]], big.mark = ",")))
  }
  cat(":::\n")
}
```

---

# Field Completeness

This table shows the completeness percentage for each output field. Fields with less than 100% completeness are highlighted.

```{r completeness-table}
if (length(report$completeness) > 0) {
  completeness_df <- data.frame(
    Field = names(report$completeness),
    Completeness = report$completeness,
    stringsAsFactors = FALSE
  )
  completeness_df <- completeness_df[order(completeness_df$Completeness), ]
  completeness_df$Status <- ifelse(
    completeness_df$Completeness >= 100, "Complete",
    ifelse(completeness_df$Completeness >= 90, "Good",
    ifelse(completeness_df$Completeness >= 70, "Fair", "Low"))
  )
  completeness_df$Completeness <- sprintf("%.1f%%", completeness_df$Completeness)

  knitr::kable(
    completeness_df,
    col.names = c("Field", "Completeness", "Status"),
    align = c("l", "r", "c"),
    row.names = FALSE
  )
} else {
  cat("No completeness data available.")
}
```

```{r completeness-chart, fig.cap="Field Completeness Distribution"}
if (length(report$completeness) > 0) {
  completeness_df <- data.frame(
    Field = names(report$completeness),
    Completeness = as.numeric(report$completeness),
    stringsAsFactors = FALSE
  )

  # Show only fields with < 100% completeness for the chart
  incomplete_df <- completeness_df[completeness_df$Completeness < 100, ]

  if (nrow(incomplete_df) > 0) {
    incomplete_df <- incomplete_df[order(incomplete_df$Completeness), ]
    incomplete_df$Field <- factor(incomplete_df$Field, levels = incomplete_df$Field)

    ggplot(incomplete_df, aes(x = Field, y = Completeness)) +
      geom_col(aes(fill = Completeness), width = 0.7) +
      geom_hline(yintercept = 90, linetype = "dashed", color = "orange", linewidth = 0.8) +
      geom_hline(yintercept = 70, linetype = "dashed", color = "red", linewidth = 0.8) +
      scale_fill_gradient2(
        low = "#dc3545",
        mid = "#ffc107",
        high = "#28a745",
        midpoint = 85,
        limits = c(0, 100)
      ) +
      coord_flip() +
      labs(
        title = "Fields with Incomplete Data",
        subtitle = "Orange line = 90%, Red line = 70%",
        x = NULL,
        y = "Completeness (%)"
      ) +
      theme_minimal() +
      theme(
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 10)
      ) +
      scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 20))
  } else {
    cat("All fields have 100% completeness.")
  }
}
```

---

# Subsection Distribution

Distribution of organizations by IRS subsection code (e.g., 501(c)(3), 501(c)(4), etc.).

```{r subsection-chart, fig.cap="Organizations by Subsection Code"}
if (!is.null(report$summary_stats$subsection_distribution)) {
  subsection_dt <- as.data.table(report$summary_stats$subsection_distribution)
  setnames(subsection_dt, c("subsection_code", "N"))

  # Calculate percentage
  subsection_dt[, pct := 100 * N / sum(N)]
  subsection_dt[, label := sprintf("%s\n(%s)",
                                    format(N, big.mark = ","),
                                    sprintf("%.1f%%", pct))]

  # Order by count descending
  subsection_dt <- subsection_dt[order(-N)]
  subsection_dt[, subsection_code := factor(subsection_code, levels = rev(subsection_code))]

  ggplot(subsection_dt, aes(x = subsection_code, y = N)) +
    geom_col(fill = "#4e79a7", width = 0.7) +
    geom_text(aes(label = format(N, big.mark = ",")),
              hjust = -0.1, size = 3.5) +
    coord_flip() +
    scale_y_continuous(
      labels = label_comma(),
      expand = expansion(mult = c(0, 0.15))
    ) +
    labs(
      title = "Organizations by Subsection Code",
      subtitle = sprintf("Total: %s organizations", format(sum(subsection_dt$N), big.mark = ",")),
      x = "Subsection Code",
      y = "Number of Organizations"
    ) +
    theme_minimal() +
    theme(
      panel.grid.major.y = element_blank(),
      axis.text.y = element_text(size = 11, face = "bold")
    )
} else {
  cat("Subsection distribution data not available.")
}
```

```{r subsection-table}
if (!is.null(report$summary_stats$subsection_distribution)) {
  subsection_dt <- as.data.table(report$summary_stats$subsection_distribution)
  setnames(subsection_dt, c("subsection_code", "N"))
  subsection_dt[, Percentage := sprintf("%.2f%%", 100 * N / sum(N))]
  subsection_dt[, N := format(N, big.mark = ",")]
  setnames(subsection_dt, c("Subsection Code", "Count", "Percentage"))

  knitr::kable(subsection_dt, align = c("l", "r", "r"))
}
```

---

# Organization Type Distribution

Distribution of organizations by exempt organization type (top 15 shown).

```{r orgtype-chart, fig.cap="Organizations by Type"}
if (!is.null(report$summary_stats$orgtype_distribution)) {
  orgtype_dt <- as.data.table(report$summary_stats$orgtype_distribution)
  setnames(orgtype_dt, c("exempt_organization_type", "N"))

  # Take top 15
  orgtype_dt <- orgtype_dt[order(-N)][1:min(15, .N)]
  orgtype_dt <- orgtype_dt[!is.na(exempt_organization_type)]

  # Truncate long names
  orgtype_dt[, display_name := ifelse(
    nchar(exempt_organization_type) > 40,
    paste0(substr(exempt_organization_type, 1, 37), "..."),
    exempt_organization_type
  )]

  # Order for plotting
  orgtype_dt[, display_name := factor(display_name, levels = rev(display_name))]

  ggplot(orgtype_dt, aes(x = display_name, y = N)) +
    geom_col(fill = "#59a14f", width = 0.7) +
    geom_text(aes(label = format(N, big.mark = ",")),
              hjust = -0.1, size = 3.5) +
    coord_flip() +
    scale_y_continuous(
      labels = label_comma(),
      expand = expansion(mult = c(0, 0.15))
    ) +
    labs(
      title = "Organizations by Exempt Organization Type",
      subtitle = "Top 15 types shown",
      x = NULL,
      y = "Number of Organizations"
    ) +
    theme_minimal() +
    theme(
      panel.grid.major.y = element_blank(),
      axis.text.y = element_text(size = 9)
    )
} else {
  cat("Organization type distribution data not available.")
}
```

```{r orgtype-table}
if (!is.null(report$summary_stats$orgtype_distribution)) {
  orgtype_dt <- as.data.table(report$summary_stats$orgtype_distribution)
  setnames(orgtype_dt, c("exempt_organization_type", "N"))
  orgtype_dt[, Percentage := sprintf("%.2f%%", 100 * N / sum(N))]
  orgtype_dt[, N := format(N, big.mark = ",")]
  setnames(orgtype_dt, c("Organization Type", "Count", "Percentage"))

  knitr::kable(head(orgtype_dt, 15), align = c("l", "r", "r"))
}
```

---

# Financial Summary

```{r financial-summary}
if (!is.null(report$summary_stats$total_assets) || !is.null(report$summary_stats$median_assets)) {
  financial_df <- data.frame(
    Metric = character(),
    Value = character(),
    stringsAsFactors = FALSE
  )

  if (!is.null(report$summary_stats$total_assets)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Total Assets",
      Value = paste0("$", format(report$summary_stats$total_assets, big.mark = ",", scientific = FALSE))
    ))
  }

  if (!is.null(report$summary_stats$median_assets)) {
    financial_df <- rbind(financial_df, data.frame(
      Metric = "Median Assets",
      Value = paste0("$", format(report$summary_stats$median_assets, big.mark = ",", scientific = FALSE))
    ))
  }

  knitr::kable(financial_df, col.names = c("Metric", "Value"), align = c("l", "r"))
} else {
  cat("Financial summary data not available.")
}
```

---

# Report Metadata

```{r metadata}
meta_df <- data.frame(
  Property = c(
    "Report Generated",
    "Pipeline Timestamp",
    "Row Preservation Check",
    "Overall Result"
  ),
  Value = c(
    format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    format(report$timestamp, "%Y-%m-%d %H:%M:%S"),
    if (report$row_preservation) "Passed" else "Failed",
    if (report$passed) "PASSED" else "FAILED"
  )
)

knitr::kable(meta_df, col.names = c("Property", "Value"), align = c("l", "l"))
```

---

*Generated by BMF Pipeline Quality System*
