---
title: "ORG Fields"
format: html
execute:
  warning: false
  message: false
  echo: false 
  output: asis
---


ORG_ fields represent stable organizational traits that typically are not changing over time. 

- EIN 
- EIN2  
- ORG_NAME 
- ORG_NAME_SEC 
- ORG_PERS_ICO 
- ORG_FISCAL_PERIOD 
- ORG_TYPE_501C 
- ORG_TYPE_PF 
- ORG_CORP_FORM 
- ORG_RULING_DATE 
- ORG_RULING_YEAR 
- ORG_YEAR_FIRST 
- ORG_YEAR_LAST 
- ORG_YEAR_COUNT 




```{r}
## Packages

library( tidyverse )   # data wrangling 
library( data.table )  # data wrangling 
library( pander )      # pretty tables 
library( knitr )       # pretty tables
library( DiagrammeR )  # drawing diagrams
library( ggthemes )    # graphics 
library( gridExtra )   # graphics

library( ggraph    )   # graphs
library( igraph    )   # graphs
library( viridis   )   # graphics


# CUSTOM FUNCTIONS: fable()
source( "R/functions.R" )

# FACTOR LABELS  
source( "data/DICTIONARIES.R" )
```


## Data from Step-00 PreProcessing

```{r, cache=T}
d <- readRDS( "data/BMF-LABELED-TEMP.rds" )
```


```{r, eval=F, echo=F}

#######
#######   VARIABLES IN THIS SECTION
#######

# "EIN", "EIN2", 
# "ORG_NAME", 
# "ORG_NAME_SEC", 
# "ORG_PERS_ICO", 
# "ORG_FISCAL_PERIOD", 
# "ORG_TYPE_501C", 
# "ORG_TYPE_PF",
# "ORG_CORP_FORM",
# "ORG_RULING_DATE", 
# "ORG_RULING_YEAR",
# "ORG_YEAR_FIRST", 
# "ORG_YEAR_LAST", 
# "ORG_YEAR_COUNT", 







# "NTEE_IRS", 
# "NTEE_NCCS",  
# "NTEEV2", 
# "NCCS_LEVEL_1", 
# "NCCS_LEVEL_2", 
# "NCCS_LEVEL_3",
# 
# "BMF_SUBSECTION_CODE",    
# "BMF_TAX_EXEMPT_PURP",
# "BMF_CLASSIFICATION_CODE",
# "BMF_DEDUCTIBILITY_CODE", 
# "BMF_FILING_REQ_CODE", 
# "BMF_PF_FILING_REQ_CODE",
# "BMF_FOUNDATION_CODE",    
# "BMF_STATUS_CODE", 
# "BMF_AFFIL_CODE",
# # "BMF_AFFILIATION_CODE",   
# "BMF_GROUP_EXEMPT_NUM",
# "BMF_INCOME_LEVEL",        
# "BMF_ASSET_LEVEL", 
# 
# "F990_FORM_YEAR",
# "F990_TOTAL_REVENUE_RECENT", 
# "F990_TOTAL_INCOME_RECENT", 
# "F990_TOTAL_ASSETS_RECENT",   
# 
# "F990_ORG_ADDR_STREET",
# "F990_ORG_ADDR_CITY", 
# "F990_ORG_ADDR_STATE", 
# "F990_ORG_ADDR_ZIP", 
# 
# "CENSUS_URBAN_AREA",
# "CENSUS_STATE_ABBR",
# "CENSUS_COUNTY_NAME",
# "CENSUS_CBSA_NAME", 
# "CENSUS_CBSA_FIPS",
# "CENSUS_BLOCK_FIPS",
#  
# "ORG_ADDR_FULL", 
# "ORG_ADDR_MATCH", 
# "LATITUDE", 
# "LONGITUDE", 
# "GEOCODER_SCORE", 
# "GEOCODER_MATCH"  
```





## Fiscal Period (ORG_FISCAL_PERIOD) 


```{r org-fiscal-period}
fable( d$ORG_FISCAL_PERIOD, p=T )
```


## Corporate Form (BMF_CORP_FORM)

```{r}
dictionaries$ORG_CORP_FORM %>% kable()
```


```{r corp-form-codes}
fable( d$ORG_CORP_FORM )
```






## Tax Exempt Type (ORG_501C_TYPE)

```{r}
fable( d$ORG_TYPE_501C, as.factor=T )
```



```{r tax-exemt-type-graph, eval=F, echo=F}
ggplot( d, aes( ORG_501C_TYPE ) ) +
     geom_bar( fill="steelblue" ) +
     theme_minimal() + 
     scale_fill_fivethirtyeight() + 
     coord_flip() +
     theme( axis.text.x=element_blank(),
            axis.title.x=element_blank() )
```




```{r 501c-circle-graph, fig.height=6, fig.width=8}
# FROM METADATA STANDARDS: 
# https://github.com/Nonprofit-Open-Data-Collective/metadata-standards/blob/main/tax-exempt-status/TAX-EXEMPT-TYPES.csv

dd <- 
structure(list(type_501c = c("501C_01", "501C_02", "501C_03", 
"501C_04", "501C_05", "501C_06", "501C_07", "501C_08", "501C_09", 
"501C_10", "501C_11", "501C_12", "501C_13", "501C_14", "501C_15", 
"501C_16", "501C_17", "501C_18", "501C_19", "501C_20", "501C_21", 
"501C_22", "501C_23", "501C_24", "501C_25", "501C_26", "501C_27", 
"501C_28", "501C_29", "501C_40", "501C_50", "501C_60", "501C_70", 
"501C_71", "501C_72", "501C_80", "501C_81"), exempt_subtype = c("CORPORATION", 
"CORPORATION", "GENERAL", "GENERAL", "SPECIAL INT GRP", "GENERAL", 
"MEMBERSHIP", "SPECIAL INT GRP", "INSURANCE", "MEMBERSHIP", "PENSION", 
"INSURANCE", "GENERAL", "COOPERATIVE", "INSURANCE", "COOPERATIVE", 
"INSURANCE", "PENSION", "MEMBERSHIP", "DEFUNCT", "INSURANCE", 
"PENSION", "SPECIAL INT GRP", "DEFUNCT", "CORPORATION", "INSURANCE", 
"INSURANCE", "PENSION", "INSURANCE", "MEMBERSHIP", "COOPERATIVE", 
"COOPERATIVE", "GENERAL", "COOPERATIVE", "GENERAL", "COOPERATIVE", 
"PENSION")), row.names = c(NA, -37L), class = "data.frame")

tt <- 
structure(list(Var1 = structure(1:40, levels = c("", "501C_00", 
"501C_01", "501C_02", "501C_03", "501C_04", "501C_05", "501C_06", 
"501C_07", "501C_08", "501C_09", "501C_10", "501C_11", "501C_12", 
"501C_13", "501C_14", "501C_15", "501C_16", "501C_17", "501C_18", 
"501C_19", "501C_20", "501C_21", "501C_23", "501C_24", "501C_25", 
"501C_26", "501C_27", "501C_29", "501C_40", "501C_50", "501C_70", 
"501C_71", "501C_80", "501C_81", "501C_82", "501C_90", "501C_91", 
"501C_92", "501C_93"), class = "factor"), Freq = c(187369L, 253L, 
800L, 9681L, 2503387L, 187435L, 91330L, 118543L, 96791L, 110634L, 
19565L, 29748L, 12L, 8219L, 13927L, 5665L, 2472L, 27L, 628L, 
6L, 57760L, 65L, 30L, 3L, 2L, 2780L, 11L, 18L, 23L, 250L, 30L, 
1L, 1L, 1383L, 1L, 29L, 790L, 1362L, 9528L, 2440L)), class = "data.frame", row.names = c(NA, 
-40L))

# add nonprofit group types
tt2 <- merge( tt, dd, by.x="Var1", by.y="type_501c" )

# setup for circle packing diagram
from     <- tt2$exempt_subtype
to       <- tt2$Var1 %>% as.character()
size     <- sqrt( tt2$Freq )  # raw counts too dramatic
edges    <- data.frame( from, to )

groups   <- unique( from )
vertices <- 
  data.frame( 
    name=c( to, groups ), 
    size=c( size,rep( 0, length(groups) ) ),
    type=c( from, paste0( groups, "_X" ) ) )

# head( vertices )
# |name    |      size|type            |
# |:-------|---------:|:---------------|
# |501C_01 |  28.28427|CORPORATION     |
# |501C_02 |  98.39207|CORPORATION     |
# |501C_04 | 432.93764|GENERAL         |
# |501C_05 | 302.20854|SPECIAL INT GRP |
# |501C_06 | 344.30074|GENERAL         |
# |501C_07 | 311.11252|MEMBERSHIP      |
  
mygraph  <- graph_from_data_frame( edges, vertices=vertices )
 
# BLACK AND WHITE VERSION 
p <- 
  ggraph(mygraph, layout = 'circlepack', weight=size) + 
  geom_node_circle() +
  theme_void()

# USE TYPE FOR SLIGHT DIFFERENCE IN 
# GROUP TYPE VERSUS MEMBERS 
# p$data$depth <- factor( p$data$type )

npgroups <- 
  c("GENERAL", "SPECIAL INT GRP", "MEMBERSHIP", "INSURANCE", 
     "PENSION", "COOPERATIVE","CORPORATION",  "DEFUNCT")

p + 
  geom_node_circle( aes( fill = as.factor(type)) ) +
  geom_node_text( aes( label=name, size=3*size, filter=leaf ), 
                  color="white", fontface="bold" ) +
  guides( size = "none") +
  scale_fill_viridis( alpha=0.5, discrete=T , 
    name="Nonprofit Function", breaks=npgroups )
```




## Ruledate (ORG_RULING_DATE)



```{r ruleyear-graph}
ggplot( d, aes( ORG_RULING_YEAR ) ) +
     geom_bar( fill="steelblue" ) +
     theme_minimal() + 
     scale_fill_fivethirtyeight() + 
     theme( axis.text.y=element_blank(),
            axis.title.y=element_blank() )
```


## Core Data Availability 

- ORG_YEAR_FIRST: first time an organization appears in an NCCS Core Database 
- ORG_YEAR_LAST: last year they appear in the NCCS Core Database 
- ORG_YEAR_COUNT: number of appearances in the NCCS Core Database 


```{r core-data-graphs, echo=F}

df <- 
  d %>% 
  select( ORG_YEAR_FIRST, ORG_YEAR_LAST ) %>% 
  gather() %>% 
  group_by( key ) %>%
  count( value )

p1 <- ggplot( d, aes( ORG_YEAR_FIRST) ) +
     geom_bar( fill="steelblue" ) +
     theme_minimal() + 
     scale_fill_fivethirtyeight() + 
     theme( axis.text.y=element_blank(),
            axis.title.y=element_blank() )

p2 <- ggplot( d, aes( ORG_YEAR_LAST ) ) +
     geom_bar( fill="steelblue" ) +
     theme_minimal() + 
     scale_fill_fivethirtyeight() + 
     theme( axis.text.y=element_blank(),
            axis.title.y=element_blank() )

p3 <- ggplot( d, aes( ORG_YEAR_COUNT) ) +
     geom_bar( fill="steelblue" ) +
     theme_minimal() + 
     scale_fill_fivethirtyeight() + 
     theme( axis.text.y=element_blank(),
            axis.title.y=element_blank() )

grid.arrange( p3, p1, p2,
  nrow = 3,
  top = "Years of Data Available in the NCCS Harmonized Core Files" )
```






<style>

h1, h2 {
  margin-top: 80px
}

.codebook-entry {
   background-color: beige;
   margin-top: 20px;
}
</style>





