---
title: "IRS Fields (BMF)"
format: html
execute:
  warning: false
  message: false
  echo: false 
  output: asis
editor: 
  markdown: 
    wrap: 72
---

Documentation for the fields that originate from the IRS Business Master
File (BMF).

```{r}
## Packages

library( tidyverse )   # data wrangling 
library( data.table )  # data wrangling 
library( pander )      # pretty tables 
library( knitr )       # pretty tables
library( DiagrammeR )  # drawing diagrams
library( ggthemes )    # graphics 
library( gridExtra )   # graphics

# CUSTOM FUNCTIONS: fable()
source( "R/functions.R" )

# FACTOR LABELS  
source( "data/DICTIONARIES.R" )
```


```{r}
library( yarrr )
library( treemap )

fx <- function( pal.type, tm.title="TITLE" ) {

  tt <- table( f ) %>% as.data.frame()
  nlevs    <- nlevels(f)
  pal      <- yarrr::piratepal( palette = pal.type )
  n.loops  <- ceiling( nlevs/length(pal) )
  step     <- 1 / n.loops 
  mix.w <- seq( from = 0.7, to = 1, length.out=(n.loops-1) )
  mix.w <- c( 0.5, mix.w )

  pal.yar <- NULL

  for( i in mix.w )
  {

    col.yar <- yarrr::piratepal( 
        palette = pal.type,
        mix.p = i )

     pal.yar <- c( pal.yar, col.yar )
  }

  tt <- dplyr::arrange( tt, -Freq )
  tt$col <- pal.yar[1:nlevs]
  tt <- dplyr::arrange( tt, f )
  tt$id <- 1:nrow(tt)

  treemap( tt,
    index="f",
    vSize="Freq",
    vColor="id",
    type="manual",
    palette=tt$col,
    border.col="gray80",
    position.legend="none",
    title=tm.title,
    algorithm="pivotSize",
    sortID="id" )

}

# f <- d$ORG_CORP_FORM
# fx( "appletv" )

# fx( "evildead" )
# fx( "up" )
# fx( "info" )
# fx( "ohbrother" )
# fx( "appletv" )
# fx( "brave" )
# fx( "espresso" )
# fx( "ipod" )
# fx( "brave" )
# fx( "decision" )
# fx( "xmen" )
# fx( "usualsuspects" )
# fx( "basel" )
# fx( "eternal" )
# fx( "google" )
# fx( "nemo" )

```

```{r, cache=T}
## Data from Step-00 PreProcessing
d <- readRDS( "data/BMF-LABELED-TEMP.rds" )
```

```{r, eval=F, echo=F}

#######
#######   VARIABLES IN THIS SECTION
#######

# "BMF_SUBSECTION_CODE",    
# "BMF_CLASSIFICATION_CODE",
# [x] "BMF_TAX_EXEMPT_PURP",
# [x] "BMF_DEDUCTIBILITY_CODE", 
# [x] "BMF_FILING_REQ_CODE", 
# [x] "BMF_PF_FILING_REQ_CODE",
# [x] "BMF_FOUNDATION_CODE",    
# [x] "BMF_STATUS_CODE", 
# [x] "BMF_AFFILIATION_CODE",   
# [x] "BMF_GROUP_EXEMPT_NUM",
# [x] "BMF_INCOME_LEVEL",        
# [x] "BMF_ASSET_LEVEL", 
 
```


<br>
<br>
<br>


## BMF_STATUS_CODE

The Exempt Organization (EO) Status Code defines the type of exemption held by the
organization. 

```{r status-codes}
fable( d$BMF_STATUS_CODE )
```


**Trust (12):**  
Trust described in section 4947(a)(2) of the Internal Revenue Code  

**Split Interest Trust (25):**  
Organization terminating its private foundation status under section 507(b)(1)(B) of the Code  

```{r, eval=F}
#| column: margin
(dictionaries$BMF_STATUS_CODE)[,c(1,3)] %>% kable( align=c("l","l") )
```





## BMF_AFFILIATION_CODE

```{dot}
digraph G {

  node [shape=plaintext]

  parent -> intermediate_east;
  parent -> intermediate_west;
  intermediate_east -> subordinate_a;
  intermediate_east -> subordinate_b;
  intermediate_west -> subordinate_c;
  intermediate_west -> subordinate_d;
}
```


Affiliation Code defines the organizational grouping. In the context of
an IRS 990 Group Exemption structures:

-   **Central Organizations**: At the top of the group exemption
    structure is the central organization, often a national or
    overarching entity that holds the group exemption letter from the
    IRS. This central organization is responsible for overseeing the
    entire group and ensuring compliance with IRS regulations.
-   **Intermediate Organizations**-: Intermediate organizations are
    entities within the group that are affiliated with the central
    organization but operate at a regional, state, or local level. These
    intermediate organizations may have their own separate tax-exempt
    status but are covered under the group exemption umbrella of the
    central organization. They typically handle operations, fundraising,
    and programs within their specific geographic area or focus.
-   **Subordinate Organizations**: Below the intermediate organizations
    are the subordinate organizations, which are the local chapters,
    branches, or affiliates of the group. These organizations are also
    covered under the group exemption and operate under the oversight of
    both the central and intermediate organizations. \`

| code | role           | affiliation      | label                        |
|-----:|:---------------|:-----------------|:-----------------------------|
|    3 | INDY           | NONE             | INDEPENDENT (3)              |
|    1 | PARENT         | NETWORK          | NETWORK-PARENT (1)           |
|    2 | INTERMEDIATE   | NETWORK          | NETWORK-INTERM (2)           |
|    6 | PARENT-REGULAR | GROUPEXM         | GROUP-EXM-PARENT-REGULAR (6) |
|    7 | INTERMEDIATE   | GROUPEXM         | GROUP-EXM-INTERM (7)         |
|    8 | PARENT-CHURCH  | GROUPEXM         | GROUP-EXM-PARENT-CHURCH (8)  |
|    9 | SUBORDINATE    | GROUPEXM/NETWORK | SUBORDINATE (9)              |


The use of intermediate organizations in a group exemption structure
allows for centralized management and oversight while still
accommodating regional or local operations and initiatives. It helps
streamline administrative processes, ensures consistency in compliance,
and provides a framework for collaboration and coordination among
affiliated entities.

```{r affiliation-codes, echo=F}
fable( d$BMF_AFFILIATION_CODE )
```


**INDEPENDENT (3)** "This code is used if the organization is an
independent organization or an independent auxiliary (i.e., not
affiliated with a National, Regional, or Geographic grouping of
organizations).",

**NETWORK-PARENT (1)** "This code is used if the organization is a
central type organization (no group exemption) of a National, Regional
or Geographic grouping of organizations.",

**NETWORK-INTERM (2)** "This code is used if the organization is an
intermediate organization (no group exemption) of a National, Regional
or Geographic grouping of organizations (such as a state headquarters of
a national organization).",

**GROUP-EXM-PARENT-REGULAR (6)** "This code is used if the organization
is a parent (group ruling) and is not a church or
501(c)(1)organization.",

**GROUP-EXM-INTERM (7)** "This code is used if the organization is a
group exemption intermediate organization of a National, Regional or
Geographic grouping of organizations.",

**GROUP-EXM-PARENT-CHURCH (8)** "This code is used if the organization
is a parent (group ruling) and is a church or 501(c)(1) organization.",

**SUBORDINATE (9)** "This code is used if the organization is a
subordinate in a group ruling.")







## BMF_DEDUCTIBILITY_CODE

Deductibility Code signifies whether contributions made to an
organization are deductible.

```{r, eval=F}
dictionaries$BMF_DEDUCTIBILITY_CODE %>% kable( align=c("l","r") )
```

```{r, eval=F}
f <- d$BMF_DEDUCTIBILITY_CODE
fx( "info", tm.title="BMF_DEDUCTIBILITY_CODE" )
```


```{r deduct-codes}
#| column: margin
fable( d$BMF_DEDUCTIBILITY_CODE )
```



## BMF_ASSET_CODE

Asset levels based upon the last year of financials reported. 

```{r, eval=F}
dictionaries$BMF_ASSET_LEVEL %>% kable( align=c("l","r") )
```

```{r asset-level, include=FALSE}
t <- fable( d$BMF_ASSET_LEVEL, as.factor=T )
```

```{r, fig.height=3}
ggplot( t, aes( x=BMF_ASSET_LEVEL, y=Freq )) + 
  geom_bar( stat = "identity", fill="steelblue" ) +
  coord_flip() +
  theme_minimal() + 
     scale_fill_fivethirtyeight() + 
     theme( axis.text.x=element_blank(),
            axis.title.y=element_blank(),
            axis.title.x=element_blank() )
```


```{r}
#| column: margin
fable( d$BMF_ASSET_LEVEL, as.factor=T )
```


## Income Levels (BMF_ASSET_CODE)

```{r, eval=F}
dictionaries$BMF_ASSET_LEVEL %>% kable( align=c("l","r") )
```


```{r income-level, include=FALSE}
#| column: margin
t <- fable( d$BMF_INCOME_LEVEL )
```

```{r, fig.height=3}
ggplot( t, aes( x=BMF_INCOME_LEVEL, y=Freq )) + 
  geom_bar( stat = "identity", fill="steelblue" ) +
  coord_flip() +
  theme_minimal() + 
     scale_fill_fivethirtyeight() + 
     theme( axis.text.x=element_blank(),
            axis.title.y=element_blank(),
            axis.title.x=element_blank() )
```

```{r}
#| column: margin
fable( d$BMF_INCOME_LEVEL )
```



## Tax Exempt Purpose Codes (BMF_TAX_EXEMPT_PURP)

When applying for tax exempt status from the IRS, nonprofit founders
report **organizational purpose** on their application, which is used by
the IRS to make a determination on whether the organization qualifies
for tax-exempt status. These codes differ from the NTEE taxonomies in
that each code is binary (yes/no), and they are NOT mutually exclusive,
so a nonprofit mission can fulfill one or several of these purposes.

IRS Tax-Exempt Purpose Codes for 501C3 Organizations: \[ See
[Instructions pp 6-7](https://www.irs.gov/pub/irs-pdf/i1023ez.pdf) \]

-   Charitable Purpose \[yes/no\]
-   Religious Purpose \[yes/no\]
-   Educational Purpose \[yes/no\]
-   Scientific Purpose \[yes/no\]
-   Literary Purpose \[yes/no\]
-   Public Safety Purpose \[yes/no\]
-   Amateur Sports Purpose \[yes/no\]
-   Prevent Cruelty to Animals and/or Children \[yes/no\]

The other 501C Types have similar tax-exempt purpose categories:

```{r, eval=F}
dictionaries$BMF_TAX_EXEMPT_PURP %>% kable( align=c("l","r") )
```

```{r tax-exempt-purp, eval=F}
fable( d$BMF_TAX_EXEMPT_PURP )
```



## Foundation Codes (BMF_FOUNDATION_CODE)

```{r, eval=F}
dictionaries$BMF_FOUNDATION_CODE %>% kable()
```

- **501C_not_3 (0):**  All organizations except 501(c)(3)
- **priv op foundation (2):** Private operating foundation exempt from paying excise taxes on
investment income
- **priv op foundation (other) (3):** Private operating foundation (other)
- **priv non-op foundation (4):** Private non-operating foundation
- **status suspended (9):** Suspense
- **church (10):** Church 170(b)(1)(A)(i)
- **school (11):** School 170(b)(1)(A)(ii)
- **hospital/ med research (12):** Hospital or medical research organization 170(b)(1)(A)(iii)
- **public university support (13):** Organization that operates for the benefit of a college or university
and is owned or operated by a governmental unit 170(b)(1)(A)(iv)
- **governmental unit (14):** Governmental unit 170(b)(1)(A)(v)
- **gov or public support (15):** Organization that receives a substantial part of its support from a
governmental unit or the general public 170(b)(1)(A)(vi)
- **one-third investment or ubi (16):** Organization that normally receives no more than one-third of its
support from gross investment income and unrelated business income and
at the same time more than one-third of its support from contributions,
fees, and gross receipts related to exempt purposes 509(a)(2)
- **subsidiary (17):** Organizations operated solely for the benefit of and in conjunction with
organizations described in 10 through 16 above 509(a)(3)
- **public safety (18):** Organization organized and operated to test for public safety 509(a)(4)
- **supporting org type1 (21):** 509(a)(3) Type I
- **supporting org type2 (22):** 509(a)(3) Type II
- **supporting org type3 (23):** 509(a)(3) Type III functionally integrated
- **supporting org type3 nfi (24):** 509(a)(3) Type III not functionally integrated



```{r foundation-codes, echo=F}
fable( d$BMF_FOUNDATION_CODE )
```


```{r}
#| column: margin
f <- d$BMF_FOUNDATION_CODE
fx( "info", tm.title="BMF_FOUNDATION_CODE" )
```





### Supporting Organization Types

**TYPE I**

A Type I supporting organization must be operated, supervised or
controlled by its supported organization(s), typically by giving the
supported organization(s) the power to regularly appoint or elect a
majority of the directors or trustees of the supporting organization.
The relationship between the supported organization(s) and the
supporting organization is sometimes described as similar to a
parent-subsidiary relationship.

**TYPE II**

A Type II supporting organization must be supervised or controlled in
connection with its supported organization(s), typically by having a
majority of the directors or trustees of the supported organization(s)
serve as a majority of the trustees or directors of the supporting
organization. The relationship between the supported organization(s) and
the supporting organization is sometimes described as similar to a
brother-sister relationship.

**TYPE III**

A Type III supporting organization must be operated in connection with
one or more publicly supported organizations. All supporting
organizations must be responsive to the needs and demands of, and must
constitute an integral part of or maintain significant involvement in,
their supported organizations. Type I and Type II supporting
organizations are deemed to accomplish these responsiveness and integral
part requirements by virtue of their control relationships. However, a
Type III supporting organization is not subject to the same level of
control by its supported organization(s). Therefore, in addition to a
notification requirement, Type III supporting organizations must pass
separate responsiveness and integral part tests.

<br>
<hr>
<br>

**FUNCTIONALLY INTEGRATED**

A Type III organization must satisfy one of three alternative integral
part tests.

**Alternative 1: Activities test**

-   Direct furtherance prong. Substantially all of the supporting
    organization’s activities must be direct furtherance activities.\
-   Direct furtherance activities are conducted by the supporting
    organization itself, rather than by a supported organization.\
-   Fundraising, managing non-exempt-use assets, grant making to
    organizations, and grant making to individuals (unless it meets
    certain requirements) are not direct furtherance activities.\
-   But for prong. In addition, substantially all of such activities
    must be activities in which, but for the supporting organization’s
    involvement, the supported organization would normally be involved.

Examples include holding and managing facilities used by a church for
its religious purposes, operating food pantries for a group of churches
that normally would operate food pantries themselves, and maintaining
local parks for a community foundation that otherwise would maintain
those parks.

**Alternative 2: Parent of supported organizations**

-   Governance. The supporting organization must have the power to
    appoint a majority of the officers, directors or trustees of each of
    its supported organizations.\
-   Substantial degree of direction. In addition, the supporting
    organization must perform “parent-like” activities by exercising a
    substantial degree of direction over the policies, programs and
    activities of the supported organizations.

**Alternative 3: Supporting a governmental entity**

-   Notice 2014-4 provides interim guidance for supporting organizations
    that want to qualify as functionally integrated because they are
    supporting a governmental entity.

Note: The transitional rule set out in the notice is not intended to
signal what future proposed regulations will require with respect to
qualifying as functionally integrated by supporting a governmental
entity. No Type III supporting organization will qualify as functionally
integrated by reason of satisfying this transitional rule once final
regulations under §509 are published.

<br>
<hr>
<br>

**NON-FUNCTIONALLY INTEGRATED**

Distribution requirement. A Type III non-FISO must distribute its
distributable amount each year to one or more of its supported
organizations.

Under final regulations, the distributable amount equals the greater of
(1) 85 percent of the organization’s adjusted net income for the prior
taxable year and (2) 3.5 percent of the aggregate fair market value of
the organization’s non-exempt use assets, with certain adjustments.

Certain excess amounts may reduce the distributable amount in subsequent
years (for up to five years after the excess amount is generated).

Attentiveness requirement. It is not enough for a Type III non-FISO to
maintain a certain level of distributions to a supported organization.
Those distributions must be sufficiently important to the supported
organization to ensure that the supported organization has sufficient
reason to pay attention to the supporting organization’s role in its
operations. Distributions to a particular supported organization are
sufficient to ensure such attentiveness if the amount of support:

Equaled at least 10 percent of the supported organization’s total
support for the supported organization’s prior year; Was necessary to
avoid interruption of the carrying on of a particular substantial
function or activity of the supported organization;

or

Based on all facts and circumstances (including actual evidence of
attentiveness) was sufficient to ensure attentiveness. At least
one-third of the supporting organization’s distributable amount must be
distributed to supported organizations (1) that meet this attentiveness
requirement and (2) to which the supporting organization is responsive.

<br>

<hr>

<br>

## Filing Requirement Code (BMF_FILING_REQ_CODE)

```{r, eval=F}
(dictionaries$BMF_FILING_REQ_CODE)[,1:2] %>% kable( align="r")
```

```{r filing-req}
fable( d$BMF_FILING_REQ_CODE )
```


```{r, fig.height=6}
#| column: margin
f <- d$BMF_FILING_REQ_CODE
fx( "info", tm.title="BMF_FILING_REQ_CODE" )
```







```{=html}
<style>

h1, h2 {
  margin-top: 80px
}

h3 {
  margin-bottom: 20px;
}

.codebook-entry {
   background-color: beige;
   margin-top: 20px;
}


.table {
   width: fit-content;
}

strong {
  font-weight: bold;
}

</style>
```
