---
title: "Lookup Tables Reference"
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(openxlsx)
library(knitr)
library(data.table)

lookup_path <- "../data/lookup/bmf_code_lookup.xlsx"
sheets <- getSheetNames(lookup_path)

# Helper function to read and display a sheet
read_lookup <- function(sheet_name) {
  df <- read.xlsx(lookup_path, sheet = sheet_name)
  setDT(df)

  # Exclude date columns (used internally for SCD Type 2)
  date_patterns <- "start_date|end_date|effective_date"
  cols_to_keep <- names(df)[!grepl(date_patterns, names(df), ignore.case = TRUE)]
  df <- df[, ..cols_to_keep]

  # Sanitize text columns to prevent markdown table breaking
  char_cols <- names(df)[sapply(df, is.character)]
  for (col in char_cols) {
    df[[col]] <- gsub("\\|", "/", df[[col]])
    df[[col]] <- gsub("\n", " ", df[[col]])
    df[[col]] <- gsub("\r", " ", df[[col]])
  }

  return(df)
}
```

This appendix contains complete definitions for all coded values in the BMF pipeline. These lookup tables are stored in `data/lookup/bmf_code_lookup.xlsx` and are used by transformation functions to map raw IRS codes to human-readable definitions.

---

# Classification Codes

## Subsection Classification Codes

Maps combinations of IRC subsection and classification codes to exempt organization types.

```{r}
#| echo: false
kable(read_lookup("subsection_classification_code"))
```

---

## Affiliation Codes

Describes an organization's relationship within a group exemption structure.

```{r}
#| echo: false
kable(read_lookup("affiliation_code"))
```

---

## Deductibility Codes

Indicates whether contributions to the organization are tax-deductible under IRC Section 170.

```{r}
#| echo: false
kable(read_lookup("deductibility_code"))
```

---

## Foundation Codes

Classifies 501(c)(3) organizations based on their public charity or private foundation status under IRC Section 509(a).

```{r}
#| echo: false
kable(read_lookup("foundation_code"))
```

---

## Organization Codes

Indicates the legal form or structure of the exempt organization.

```{r}
#| echo: false
kable(read_lookup("organization_code"))
```

---

## Status Codes

Indicates the current standing of an organization's tax-exempt recognition with the IRS.

```{r}
#| echo: false
kable(read_lookup("status_code"))
```

---

# Activity Codes

## IRS Activity Codes

3-digit codes describing the primary purposes or activities of an exempt organization. Each organization may report up to three activity codes.

```{r}
#| echo: false
kable(read_lookup("activity_code"))
```

---

# NTEE Codes

## NTEE Code Definitions

The National Taxonomy of Exempt Entities (NTEE) classification system codes and definitions.

```{r}
#| echo: false
kable(read_lookup("ntee_code"))
```

---

## NTEE Major Groups

Top-level NTEE categories identified by the first letter of the NTEE code.

```{r}
#| echo: false
kable(read_lookup("ntee_code_major_group"))
```

---

## NTEE Common Codes

Common codes that can appear across multiple NTEE major groups.

```{r}
#| echo: false
kable(read_lookup("ntee_common_code"))
```

---

# Financial Codes

## Asset Codes

Single-digit codes (0-9) placing organizations into size categories based on total assets.

```{r}
#| echo: false
kable(read_lookup("asset_code"))
```

---

## Income Codes

Single-digit codes (0-9) placing organizations into size categories based on total income.

```{r}
#| echo: false
kable(read_lookup("income_code"))
```

---

# Filing Requirement Codes

## Filing Requirement Codes

Indicates which annual information return the organization is required to file.

```{r}
#| echo: false
kable(read_lookup("filing_requirement_code"))
```

---

## Private Foundation Filing Requirement Codes

Filing requirements specific to private foundations (Form 990-PF).

```{r}
#| echo: false
kable(read_lookup("pf_filing_requirement_code"))
```

---

# Organization Name Codes

## Parent Organization Lookup

Maps abbreviations and Group Exemption Numbers to parent organization names.

```{r}
#| echo: false
kable(read_lookup("parent_organization"))
```
